---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)

shs_agg_resp <- read_csv("raw_data/shs_aggregate_responses.csv")

shs_agg_resp %>% 
  filter(!is.na(nearest_green_space_use))

shs_agg_resp %>% 
  group_by(year) %>% 
  summarise(n = n())

shs_agg_resp %>% 
  distinct(volunteering_last_twelve_months)

glimpse(shs_agg_resp)

```

Q. How do people in neighbourhoods with good access to green space differ from those who have no good access?

How do people in neighbourhoods with [green space == "A 5 minute walk or less"] differ from those who have [green space > A 10 minute walk], i.e. What are the variable combinations that predict these opposite outcomes?

Explanatory model

```{r}
shs_local_green_space_access <- shs_agg_resp %>% 
  mutate(local_access_to_green = if_else(distance_to_nearest_green_space == "A 5 minute walk or less", "Yes", "No"))

```


Q. Are there differences in how they rate their neighbourhoods? 

For a given value of [access to green space: local or not local] how do people rate thir neighbourhood?

H0 - there is no difference in how people rate their neighbourhood varying with local access to green space?

Q. Are there differences in how they rate their communities?

H0 - there is no difference in how people rate their neighbourhood varying with local access to green space?

